---
title: Prediction of flight onset of *Cacopsylla melanoneura* and *C. picta*, vectors
  of apple proliferation disease, using a temperature-based immigration analysis -
  Appendix A
author: "Bernd Panassiti, Nicolas Sander"
date: 'created: 25.10.2016, last modified: `r format(Sys.Date(), format="%d.%m.%Y")`'
output:
  md_document:
    variant: markdown_github
  html_document:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
header-includes: \usepackage{graphicx}
---


```{r setup, echo=FALSE,eval=TRUE,warning=FALSE,message=FALSE}
rm(list=ls(all=TRUE))
library(knitr) # for creating html-file
knitr::opts_knit$set(root.dir='../') # definining working directory; or normalizePath('../')
knitr::opts_chunk$set(fig.align='center') # aligns all figures
knitr::opts_chunk$set(echo=TRUE) # shows r-code
knitr::opts_chunk$set(message=FALSE) # suppresses library outputs
knitr::opts_chunk$set(warning=FALSE) # suppresses library warnings
```


# Prediction of flight onset of *Cacopsylla melanoneura* and *C. picta*, vectors
  of apple proliferation disease, using a temperature-based immigration analysis -
  Appendix A

## Introduction
This document is an appendix to the article: ''Prediction of flight onset of *Cacopsylla melanoneura* and *C. picta*, vectors of apple proliferation disease, using a temperature-based immigration analysis''.
The document provides the calculation procedure for a temperature-based immigation analysis based on Tedeschi et al. (2012).



## Data
##### Loading data and libraries

```{r load_data}
source("r-code/00_settings.R")
detachAllPackages() #detach("package:plyr", unload=TRUE) # makes problems with the group_by function
load(file="data/AP_rawdata_vector.RData")
load(file="data/AP_workingdata_clustering.RData")
load(file="data/AP_workingdata_vector.RData")
set.seed(2015)
```

```{r load_libraries}
if(!require(ggplot2))
{
  install.packages("ggplot2", repos = mir)
  require(ggplot2)
}
if(!require(plyr))
{
  install.packages("plyr")
  require(plyr)
}
if(!require(dplyr))
{
  install.packages("dplyr")
  require(dplyr)
}

if(!require(stringi))
{
  install.packages("stringi", repos = mir)
  require(stringi)
}

if(!require(scales))
{
  install.packages("scales", repos = mir)
  require(scales)
}
if(!require(MASS))
{
  install.packages("MASS", repos = mir)
  require(MASS)
}
if(!require(qtlcharts))
{
  install.packages("qtlcharts", repos = mir)
  require(qtlcharts)
}
```

#### Data preparation

P = parents, re-emigrants, overwintering vectors  
F1 = offspring, emigrants  
B = vector sampling method: beating tray  
YT = vector sampling method: yellow traps

```{r data_preperation,results="hide"}
tempsum <- data.frame()

#pulling out site_id,,cluster, year, date and abundance for each insect_id and insect_stage_id
c.melPB <- c.melanoneuraPB[,c(2,3,4,5,6,10,11)]
c.melF1B <- c.melanoneuraF1B[,c(2,3,4,5,6,10,11)]
c.picPB <- c.pictaPB[,c(2,3,4,5,6,10,11)]
c.picF1B <- c.pictaF1B[,c(2,3,4,5,6,10,11)]
c.melPYT <- c.melanoneuraPYT[,c(2,3,4,5,6,10,11)]
c.melF1YT <- c.melanoneuraF1YT[,c(2,3,4,5,6,10,11)]
c.picPYT <- c.pictaPYT[,c(2,3,4,5,6,10,11)]
c.picF1YT <- c.pictaF1YT[,c(2,3,4,5,6,10,11)]

#renaming 'amount'
names(c.melPB)[6] <- 'abundance'
names(c.melF1B)[6] <- 'abundance'
names(c.picPB)[6] <- 'abundance'
names(c.picF1B)[6] <- 'abundance'
names(c.melPYT)[6] <- 'abundance'
names(c.melF1YT)[6] <- 'abundance'
names(c.picPYT)[6] <- 'abundance'
names(c.picF1YT)[6] <- 'abundance'

clusterSites <- unique(modeldata.t[,c(2,3)]) #site_id and cluster id
regionSites <- unique(sitesInsectVector[,c(1,3)])

#Combining beating and yellow trap subsets
c.melPcomb <- merge(c.melPB, c.melPYT, by = c("site_id","region","date","cluster","year"), all = TRUE)
c.melPcomb <- transform(c.melPcomb, abundance = rowSums(c.melPcomb[,c("abundance.x","abundance.y")], na.rm = TRUE))
c.melPcomb[, "presence"] <- apply(c.melPcomb[, c("presence.x","presence.y")], 1, max, na.rm=T)
c.melPcomb <- c.melPcomb[,-which(names(c.melPcomb) %in% c("abundance.x","abundance.y","presence.x","presence.y"))]#dropping uncombined columns
###
c.melF1comb <- merge(c.melF1B, c.melF1YT, by = c("site_id","region","date","cluster","year"), all = TRUE)
c.melF1comb <- transform(c.melF1comb, abundance = rowSums(c.melF1comb[,c("abundance.x","abundance.y")], na.rm = TRUE))
c.melF1comb[, "presence"] <- apply(c.melF1comb[, c("presence.x","presence.y")], 1, max, na.rm=T)
c.melF1comb <- c.melF1comb[,-which(names(c.melF1comb) %in% c("abundance.x","abundance.y","presence.x","presence.y"))]#dropping uncombined columns
###
c.picPcomb <- merge(c.picPB, c.picPYT, by = c("site_id","region","date","cluster","year"), all = TRUE)
c.picPcomb <- transform(c.picPcomb, abundance = rowSums(c.picPcomb[,c("abundance.x","abundance.y")], na.rm = TRUE))
c.picPcomb[, "presence"] <- apply(c.picPcomb[, c("presence.x","presence.y")], 1, max, na.rm=T)
c.picPcomb <- c.picPcomb[,-which(names(c.picPcomb) %in% c("abundance.x","abundance.y","presence.x","presence.y"))]#dropping uncombined columns
###
c.picF1comb <- merge(c.picF1B, c.picF1YT, by = c("site_id","region","date","cluster","year"), all = TRUE)
c.picF1comb <- transform(c.picF1comb, abundance = rowSums(c.picF1comb[,c("abundance.x","abundance.y")], na.rm = TRUE))
c.picF1comb[, "presence"] <- apply(c.picF1comb[, c("presence.x","presence.y")], 1, max, na.rm=T)
c.picF1comb <- c.picF1comb[,-which(names(c.picF1comb) %in% c("abundance.x","abundance.y","presence.x","presence.y"))]#dropping uncombined columns
###

#TEMPSUMLIST CREATION
#listnames <- c('c.melP','c.melF1','c.picP','c.picF1','c.melPYT','c.melF1YT','c.picPYT','c.picF1YT')
#tempsumlist <- list(c.melP,c.melF1,c.picP,c.picF1,c.melPYT,c.melF1YT,c.picPYT,c.picF1YT)
listnames <- c('c.melPcomb','c.melF1comb','c.picPcomb','c.picF1comb')
tempsumlist <- list(c.melPcomb,c.melF1comb,c.picPcomb,c.picF1comb)

names(tempsumlist) <- listnames


###############DECIDE IF SUM BY REGION OR CLUSTER HERE! #########################
#Sum by region
for(i in 1:length(tempsumlist)){
tempsumlist[[i]] <- as.data.frame(tempsumlist[[i]] %>% dplyr::group_by(date,year,region) %>% dplyr::summarise(abundance = sum(abundance,na.rm=T), presence=max(presence))) #combine all sites within same region at same date
}

list2env(tempsumlist,environment())  #overwrite single dataframes in the global environment

#merge all dataframes by columns
tempsum <- join_all(list(tempsum, c.melPcomb,c.melF1comb,c.picPcomb,c.picF1comb), type="full")
```



## Temperature-based immigration analysis

##### Calculation of thresholds and indices

1) a0 and aMax
```{r a0 and aMax}
##################################################################################################
#CALCULATING A0's = Date of first occurence for each site and year (+ species, generation and method)
a0 <-lapply(tempsumlist, function(j){
  sapply(split(j, list(j[,3],j[,2])), function(x){x[,1][which(x[,5]>0)[1]]}) #split by region||cluster||site + year and look for first date where insects are found
})
a0 <- lapply(a0, function(x) x[!is.na(x)]) #removing sites and year where no insect was found thus returning NA
for(i in 1:length(a0)){
  a0[[i]] <- as.data.frame(a0[[i]])
  a0[[i]]$datetime <- as.POSIXct(a0[[i]][,1], origin="1970-01-01")
  a0[[i]]$year <- format(a0[[i]]$datetime, format="%Y")
  #a0[[i]]$t0maxmin <- NA
  #a0[[i]]$t0maxmean <- NA
}

for(i in 1:length(a0)){ #Adding region||cluster||site_id to a0 as id
  for(j in 1:length(a0[[i]][,1])){
    a0[[i]]$region <- rownames(a0[[i]][1])
    a0[[i]]$region <- gsub('.{5}$', '', a0[[i]]$region) #cut off last 5 characters (e.g. Laimburg0001.2016 -> Laimburg0001 || Bozen.2013 <- Bozen for easier filtering of overall site min t0max)
  }
}

##################################################################################################
#CALCULATING aMax's = Date of max. abundance for each site (+ species, generation and method)
aMax <-lapply(tempsumlist, function(j){
  sapply(split(j, list(j[,3])), function(x){x[,1][which.max(x[,5])]}) #split by region||cluster||site and look for first date where max # of insects are found
})
aMax <- lapply(aMax, function(x) x[!is.na(x)])
for(i in 1:length(aMax)){
  aMax[[i]] <- as.data.frame(aMax[[i]])
  aMax[[i]]$datetime <- as.POSIXct(aMax[[i]][,1], origin="1970-01-01")
  aMax[[i]]$year <- format(aMax[[i]]$datetime, format="%Y")
}

for(i in 1:length(aMax)){ #Adding region||cluster||site_id to aMax as id
  for(j in 1:length(aMax[[i]][,1])){
    aMax[[i]]$region <- rownames(aMax[[i]][1])
  }
}
```

```{r thourly & thourlyregion}
##################################################################################################
thourly <- weatherStationsHourly_split
thourly <- lapply(thourly, function(x) x[,-c(6:11)]) #removing unneeded columns
for(i in 1:nrow(clusterStations)){
  thourly[[i]]$cluster <- clusterStations$group[i]
}

for(i in 1:length(thourly)){
  thourly[[i]]$datetimenum <- as.POSIXct(thourly[[i]]$datetime, format="%Y-%m-%d")#cut off %H:%M:%S due to problems with numeric transformation
  thourly[[i]]$datetimenum <- as.numeric(thourly[[i]]$datetimenum)
}

#################################################
thourlyregion <- dplyr::bind_rows(thourly) #combine lists to 1 dataframe for dplyr
thourlyregion <- as.data.frame(thourlyregion %>% dplyr::group_by(datetime,datetimenum,region) %>% dplyr::summarise(temphourmean=mean(temphour, na.rm=T)))
thourlyregion <- split(thourlyregion, thourlyregion$region) #combine dataframes back into list

```


2) Tmax and Tmedian
```{r tmax and tmedian}
##################################################################################################
#CALCULATING TMax = Max. daily temperature & tmedian from hourly mean temps
tmax <- list()
tmedian <- list()


for(i in 1:length(weatherStationsHourly_split)){
tmax[[i]] <- aggregate(list(maxtempday=weatherStationsHourly_split[[i]][,4]),list(datetime=cut(as.POSIXct(weatherStationsHourly_split[[i]][,3]), "day")),FUN=max,na.rm=T)
tmax[[i]]$maxtempday[is.infinite(tmax[[i]]$maxtempday)] <- NA
tmax[[i]]$datetime[is.infinite(tmax[[i]]$datetime)] <- NA
tmax[[i]]$site_id <- names(weatherStationsHourly_split)[[i]]
tmax[[i]]$region  <- unique(weatherStationsHourly_split[[i]][1,5]) #repeating the stations region for easier filtering

tmedian[[i]] <- aggregate(list(mediantempday=weatherStationsHourly_split[[i]][,4]),list(datetime=cut(as.POSIXct(weatherStationsHourly_split[[i]][,3]), "day")),FUN=median,na.rm=T)
tmedian[[i]]$mediantempday[is.infinite(tmedian[[i]]$mediantempday)] <- NA
tmedian[[i]]$datetime[is.infinite(tmedian[[i]]$datetime)] <- NA
tmedian[[i]]$site_id <- names(weatherStationsHourly_split)[[i]]
tmedian[[i]]$region  <- unique(weatherStationsHourly_split[[i]][1,5]) #repeating the stations region for easier filtering
}


#naming the stations
names(tmax)    <- names(weatherStationsHourly_split)
names(tmedian) <- names(weatherStationsHourly_split)

#changing the datetime entry to numeric for easier filtering
for(i in 1:length(tmax)){
  tmax[[i]]$datetime <- as.POSIXct(tmax[[i]]$datetime, format="%Y-%m-%d")
  tmax[[i]]$datetime <- as.numeric(tmax[[i]]$datetime)
}

for(i in 1:length(tmedian)){
  tmedian[[i]]$datetime <- as.POSIXct(tmedian[[i]]$datetime, format="%Y-%m-%d")
  tmedian[[i]]$datetime <- as.numeric(tmedian[[i]]$datetime)
}


##################################################################################################
#adding cluster id & region to tmax and tmedian
for(i in 1:nrow(clusterStations)){
  tmax[[i]]$cluster <- clusterStations$group[i]
  tmedian[[i]]$cluster <- clusterStations$group[i]
}
```


3) T0max
```{r t0max,results="hide"}
##################################################################################################
#T0Max = max of max daily temperature of 7 days preceding a0 (per observation, insect type and site)
#check a0 and then filter through txmax and tmedian at corresponding weatherstation (from cluster)
#using tmax
for(i in 1:length(a0)){ #loop through all species,generation and method
  for(j in 1:length(a0[[i]][,1])){ #loop through the a0 entries
    tmaxvec <- numeric()
    for(x in 1:length(tmax)){ #loop through ALL weather stations
      if(tmax[[x]]$region[1] == a0[[i]]$region[j] ){ #check which stations cluster id matches the current sites' cluster id
  tmaxvec  <- c(tmaxvec,max(tmax[[x]]$maxtempday[(which(tmax[[x]]$datetime == a0[[i]][,1][j])-7):(which(tmax[[x]]$datetime == a0[[i]][,1][j]))])) #temporary vector saving the max of daily max temperatures of the 7 days before a0 from all weather stations x that share the id with the site j
       print(paste(i,j,x))
      }
    }
    print(tmaxvec)
    a0[[i]]$t0max[j] <- min(tmaxvec, na.rm=T) #min of all weather station's 7 day max
    a0[[i]]$t0mean[j] <- mean(tmaxvec, na.rm=T) #mean of all max weekly temperatures
  }
 a0[[i]]$t0max[is.infinite(a0[[i]]$t0max)] <- NA
 a0[[i]]$t0mean[is.nan(a0[[i]]$t0mean)] <- NA
}
```


4) T0maxmin and T0maxmean
```{r t0maxmin and t0maxmean}
##################################################################################################
#Minimum of all t0max/t0mean that belong to the same site over several years
t0maxmin <- list()
for(i in 1:length(a0)){
t0maxmin[[i]] <- as.data.frame(a0[[i]] %>% dplyr::group_by(region) %>% dplyr::summarise(t0maxmin = min(t0max,na.rm=T), t0meanmin = min(t0mean,na.rm=T)))
}
names(t0maxmin) <- names(a0)

for(i in 1:length(a0)){
  a0[[i]] <- join_all(list(a0[[i]], t0maxmin[[i]]), by="region", type="left")
}
```


5) T0maxyears
```{r t0maxyears}
##################################################################################################
#Min of t0maxmin by year for id (region || cluster) in order to check for regional differences
t0maxyears <- list()
for(i in 1:length(a0)){
t0maxyears[[i]] <- as.data.frame(a0[[i]] %>% dplyr::group_by(region) %>% dplyr::summarise(t0maxyears= min(t0maxmin,na.rm=T)))
}
names(t0maxyears) <- names(a0)

for(i in 1:length(a0)){
  a0[[i]] <- join_all(list(a0[[i]], t0maxyears[[i]]), by="region", type="left")
}
```



6) DD
```{r DD for only a0,results="hide"}
##################################################################################################
#Degree days DD = number of hours over t0max per week preceding a0

for(i in 1:length(a0)){
  for(j in 1:length(a0[[i]][,1])){
    DD <- numeric()
      for(x in 1:length(thourly)){
       #if(thourly[[x]]$cluster[1] == a0[[i]]$cluster[j]){
        if(thourly[[x]]$region[1] == a0[[i]]$region[j]){
          DD <-  c(DD,sum(as.numeric(thourly[[x]]$temphour[(which(thourly[[x]]$datetimenum == a0[[i]][,1][j])-(24*7)):(which(thourly[[x]]$datetimenum == a0[[i]][,1][j]))])>a0[[i]]$t0maxmin[j])) #(HINT: 24 hours * 7 days)
       }
      }
    print(DD)
    a0[[i]]$DD[j] <- mean(DD, na.rm=T)
  }
}
```



7) T7n
```{r T7n for only a0}
############################################################################################
# mean max daily temperatures in 7 days before a0 (using Tmax (max daily temperatures))
for(i in 1:length(a0)){
  for(j in 1:length(a0[[i]][,1])){
    T7n <- numeric()
      for(x in 1:length(tmax)){
       #if(tmax[[x]]$cluster[1] == a0[[i]]$cluster[j]){
        if(thourly[[x]]$region[1] == a0[[i]]$region[j]){
          T7n <-  c(T7n,tmax[[x]]$maxtempday[(which(tmax[[x]]$datetime == a0[[i]][,1][j])-(7)):(which(tmax[[x]]$datetime == a0[[i]][,1][j]))]) #(HINT: 7 days since daily temp data)
       }
      }
    print(T7n)
    a0[[i]]$T7n[j] <- mean(T7n, na.rm=T)
  }
}
```


8) Ii
```{r Ii for only a0}
##################################################################################################
#Immigration Index for the a0 
#Ii = (T7n - T7th) + DD
#   = mean of max in 7 days before a0 (T7n) - absolute highest temp in week before a0 (t0max) + hours per week above t0max (DD)
for(i in 1:length(a0)){
  for(j in 1:length(a0[[i]][,1])){
    a0[[i]]$Ii[j] <- (a0[[i]]$T7n[j] - a0[[i]]$t0max[j]) + a0[[i]]$DD[j]
  }
}
```

```{r ENTERING DESIRED TEMPSUMLIST INDEX}
#Index of tempsumlist of dataset you want to look at #########################################
                                          x<-1
##############################################################################################
```

```{r tmax, DD, Ii for thourly}
#Calculating max temp per day (tmax) and hours over t0max per day (DD)
for(i in 1:length(thourly)){
  thourly[[i]]$DD <- NA
  thourly[[i]]$tmax <- NA
  for(j in 24:length(thourly[[i]][,3])){
    for(x in x){  #just for one dataset (=x) of tempsumlist
      for(y in 1:length(t0maxyears[[x]][,1])){
      if(thourly[[i]]$region[j] == t0maxyears[[x]]$region[y]){
      thourly[[i]]$DD[(j-24):j] <- sum(as.numeric(thourly[[i]]$temphour[(j-(24)):j])>t0maxyears[[x]]$t0maxyears[y])
      thourly[[i]]$tmax[(j-24):j] <- max(thourly[[i]]$temphour[(j-(24)):j])
        }
      }
    }
  }
}
```

```{r Ii region}
#Ii aggregated for each id (taking means of all stations within the same id (region || cluster))
thourlyregion <- dplyr::bind_rows(thourly) #combine lists to 1 dataframe for dplyr
thourlyregion <- as.data.frame(thourlyregion %>% dplyr::group_by(datetime,datetimenum,region) %>% dplyr::summarise(temphourmean=mean(temphour, na.rm=T), ddmean=mean(DD,na.rm=T), tmaxmean=mean(tmax, na.rm=T)))

names(t0maxyears[[x]]) <- c("region", "t0maxyears") #rename for use in join_all
thourlyregion <- join_all(list(thourlyregion, t0maxyears[[x]]), by="region", type="left")
thourlyregion$Ii <- NA

thourlyregion <- split(thourlyregion, thourlyregion$region) #combine dataframes back into list


for(i in 1:length(thourlyregion)){
  for(j in 168:length(thourlyregion[[i]][,1])){
  thourlyregion[[i]]$Ii[(j-(24*7)):j] <- (mean(thourlyregion[[i]]$tmaxmean[(j-(24*7)):j]) - thourlyregion[[i]]$t0maxyears[j]) + mean((thourlyregion[[i]]$ddmean[(j-(24*7)):j])>thourlyregion[[i]]$t0maxyears[j])
  }
}
```

```{r tmax, DD, Ii region}
thourlyregion <- dplyr::bind_rows(thourlyregion) #combine lists to 1 dataframe for dplyr
names(t0maxyears[[x]]) <- c("region", "t0maxyears") #rename for use in join_all
thourlyregion <- join_all(list(thourlyregion, t0maxyears[[x]]), by="region", type="left")
thourlyregion <- split(thourlyregion, thourlyregion$region) #combine dataframes back into list

for(i in 1:length(thourlyregion)){
  thourlyregion[[i]]$DD <- NA
  thourlyregion[[i]]$tmax <- NA
  for(j in 24:length(thourlyregion[[i]][,1])){
    thourlyregion[[i]]$DD[(j-24):j] <- sum(as.numeric(thourlyregion[[i]]$temphourmean[(j-(24)):j])>thourlyregion[[i]]$t0maxyears[j])
      thourlyregion[[i]]$tmax[(j-24):j] <- max(thourlyregion[[i]]$temphourmean[(j-(24)):j])
  }
}

for(i in 1:length(thourlyregion)){
  thourlyregion[[i]]$Ii <- NA
  for(j in 168:length(thourlyregion[[i]][,1])){
  thourlyregion[[i]]$Ii[(j-(24*7)):j] <- (mean(thourlyregion[[i]]$tmax[(j-(24*7)):j]) - thourlyregion[[i]]$t0maxyears[j]) + mean((thourlyregion[[i]]$DD[(j-(24*7)):j])>thourlyregion[[i]]$t0maxyears[j])
  }
}
```



#### Graphs

```{r base plots, echo=TRUE}
Sys.setlocale("LC_TIME", "English") # to match English date structure
#Region
for(j in x){
###REGION 1-3
par(oma=c(5.5,3,2,2))
par(mfrow=c(3,1))
for(i in 1:3){
par(mar=c(5,5,4,5)+.1)
plot(thourlyregion[[i]]$datetime, thourlyregion[[i]]$Ii, main=names(thourlyregion[i]), type="l", ylab="Immigration Index (Ii)",xlab="",yaxt="n", ylim=c(-15,40),col="darkgrey",xaxt="n",cex.lab=2,cex.main=2, cex.axis=2)
abline(h=0, col="darkgrey")

for(y in c("2013","2014","2015","2016")){
tmp <- subset(thourlyregion[[i]], format(datetime, "%Y") %in% c(y))
abline(v=as.POSIXct(tmp$datetime[which(tmp$Ii > 0)[1]],format="%Y-%m-%d %H:%M%S"), col="darkgrey",lty=2,lwd=4)
abline(v=as.POSIXct(tmp$datetime[which(tmp$temphourmean >= thourlyregion[[i]]$t0maxyears[1])[1]],format="%Y-%m-%d %H:%M%S"), col=adjustcolor("red",0.3),lty=2,lwd=4)
}

points(subset(tempsumlist[[j]], region==names(thourlyregion)[i])$date, subset(tempsumlist[[j]], region==names(thourlyregion)[i])$presence * 40,col = adjustcolor(ifelse(subset(tempsumlist[[j]], region==names(thourlyregion)[i])$presence > 0,'blue','red'),0.5), pch=16,xaxt="n",cex=2)
text(as.POSIXct("2016-06-01"), thourlyregion[[i]]$t0maxyears[1]+5, "T7th", col = "red",cex=3) 
text(as.POSIXct("2013-02-01"), 0+5, "Ii=0", col = "black",cex=3) 
text(as.POSIXct("2014-11-01"), 35, paste("T7th= ",round(thourlyregion[[i]]$t0maxyears[1],digits=2)," °C"),cex=2)
####x axis###
minor <- seq(as.POSIXct("2013-01-01", format = "%Y-%m-%d"), tail(thourlyregion[[i]]$datetime, 1),
             by = "1 months")
labDates <- seq(as.POSIXct("2013-01-01", format = "%Y-%m-%d"),tail(thourlyregion[[i]]$datetime,1), by = "2 months")
axis.POSIXct(side = 1, thourlyregion[[i]]$datetime, at = labDates, format = "%b", las = 2,cex.axis=2)
axis.POSIXct(side = 1, thourlyregion[[i]]$datetime, at = minor, labels = FALSE, tcl = -0.25)
axis.POSIXct(side = 1, thourlyregion[[i]]$datetime, format = "%Y",las=1, line=4,lty=0,cex.axis=2)
###y temperature###
lines(thourlyregion[[i]]$datetime, thourlyregion[[i]]$temphourmean,yaxt="n",xaxt="n", ylab="",xlab="",col=adjustcolor("red",0.3), type="l", ylim=c(-15,40))
axis(4, cex.axis=2)
abline(h=round(thourlyregion[[i]]$t0maxyears[1],digits=2), col="red")
mtext("Mean Hourly Temperature °C",side=4,line=3,cex=1)
}
title("C. melanoneura P-generation", outer=T,cex.main=2)
#title(names(tempsumlist)[j], outer=T,cex.main=2)
###legend###
par(fig = c(0, 1, 0, 1), oma = c(0, 0, 0, 0), mar = c(0, 0, 0, 0), new = TRUE)
plot(0, 0, type = "n", bty = "n", xaxt = "n", yaxt = "n")
legend("bottom", c("Ii","T[°C]","1st Ii>0","1st T>T7th","Presence", "Absence"), col=c("darkgrey","red","darkgrey","red","blue","red"),lty=c(1,1,2,2,NA,NA),pch=c(NA,NA,NA,NA,16,16),bty="n",cex=1.75,horiz = T,pt.cex=2,text.width = 0.21, seg.len=0.75,lwd=2,xpd=2)

###REGION 4-6
par(oma=c(5.5,3,2,2))
par(mfrow=c(3,1))
for(i in 4:6){
par(mar=c(5,5,4,5)+.1)
plot(thourlyregion[[i]]$datetime, thourlyregion[[i]]$Ii, main=names(thourlyregion[i]), type="l", ylab="Immigration Index (Ii)",xlab="",yaxt="n", ylim=c(-15,40),col="darkgrey",xaxt="n",cex.lab=2,cex.main=2, cex.axis=2)
abline(h=0, col="darkgrey")

for(y in c("2013","2014","2015","2016")){
tmp <- subset(thourlyregion[[i]], format(datetime, "%Y") %in% c(y))
abline(v=as.POSIXct(tmp$datetime[which(tmp$Ii > 0)[1]],format="%Y-%m-%d %H:%M%S"), col="darkgrey",lty=2,lwd=4)
abline(v=as.POSIXct(tmp$datetime[which(tmp$temphourmean >= thourlyregion[[i]]$t0maxyears[1])[1]],format="%Y-%m-%d %H:%M%S"), col=adjustcolor("red",0.3),lty=2,lwd=4)
}

points(subset(tempsumlist[[j]], region==names(thourlyregion)[i])$date, subset(tempsumlist[[j]], region==names(thourlyregion)[i])$presence * 40,col = adjustcolor(ifelse(subset(tempsumlist[[j]], region==names(thourlyregion)[i])$presence > 0,'blue','red'),0.5), pch=16,xaxt="n",cex=2)
text(as.POSIXct("2016-06-01"), thourlyregion[[i]]$t0maxyears[1]+5, "T7th", col = "red",cex=3) 
text(as.POSIXct("2013-02-01"), 0+5, "Ii=0", col = "black",cex=3) 
text(as.POSIXct("2014-11-01"), 35, paste("T7th= ",round(thourlyregion[[i]]$t0maxyears[1],digits=2)," °C"),cex=2)
####x axis###
minor <- seq(as.POSIXct("2013-01-01", format = "%Y-%m-%d"), tail(thourlyregion[[i]]$datetime, 1),
             by = "1 months")
labDates <- seq(as.POSIXct("2013-01-01", format = "%Y-%m-%d"),tail(thourlyregion[[i]]$datetime,1), by = "2 months")
axis.POSIXct(side = 1, thourlyregion[[i]]$datetime, at = labDates, format = "%b", las = 2,cex.axis=2)
axis.POSIXct(side = 1, thourlyregion[[i]]$datetime, at = minor, labels = FALSE, tcl = -0.25)
axis.POSIXct(side = 1, thourlyregion[[i]]$datetime, format = "%Y",las=1, line=4,lty=0,cex.axis=2)
###y temperature###
lines(thourlyregion[[i]]$datetime, thourlyregion[[i]]$temphourmean,yaxt="n",xaxt="n", ylab="",xlab="",col=adjustcolor("red",0.3), type="l", ylim=c(-15,40))
axis(4, cex.axis=2)
abline(h=round(thourlyregion[[i]]$t0maxyears[1],digits=2), col="red")
mtext("Mean Hourly Temperature °C",side=4,line=3,cex=1)
}
title("C. melanoneura P-generation", outer=T,cex.main=2)
#title(names(tempsumlist)[j], outer=T,cex.main=2)
###legend###
par(fig = c(0, 1, 0, 1), oma = c(0, 0, 0, 0), mar = c(0, 0, 0, 0), new = TRUE)
plot(0, 0, type = "n", bty = "n", xaxt = "n", yaxt = "n")
legend("bottom", c("Ii","T[°C]","1st Ii>0","1st T>T7th","Presence", "Absence"), col=c("darkgrey","red","darkgrey","red","blue","red"),lty=c(1,1,2,2,NA,NA),pch=c(NA,NA,NA,NA,16,16),bty="n",cex=1.75,horiz = T,pt.cex=2,text.width = 0.21, seg.len=0.75,lwd=2,xpd=2)

}
```




```{r ggplot2}
library(gridExtra)
library(grid)
library(scales)

#Region  ##########################################################################
pl <- lapply(1:6, function(i)
ggplot()+  
geom_line(data=thourlyregion[[i]],aes(x=thourlyregion[[i]]$datetime,y=thourlyregion[[i]]$Ii),size=0.5, alpha=0.5) + 

geom_line(data=thourlyregion[[i]],aes(x=thourlyregion[[i]]$datetime,y=thourlyregion[[i]]$temphourmean),size=0.5, alpha=0.5, col="red") +
  
ylim(-10,40) + 

geom_point(data=subset(tempsumlist[[x]], region==names(thourlyregion)[i]),
           aes(x=subset(tempsumlist[[x]], region==names(thourlyregion)[i])$date, y=subset(tempsumlist[[x]], region==names(thourlyregion)[i])$abundance,color=subset(tempsumlist[[x]], region==names(thourlyregion)[i])$abundance > 0),alpha=0.3) +

scale_colour_manual(name = '+ Presence', values = setNames(c('darkblue','red'),c(T, F)),guide=FALSE) +
geom_hline(yintercept = 0, col="blue") +
geom_hline(yintercept = round(thourlyregion[[i]]$t0maxyears[1],digits=2), col="red") +

xlab("Date") +
ylab("Immigration Index") +
theme_light() + 
ggtitle(names(thourlyregion[i])) +
annotate(geom="text", x=as.POSIXct("2016-03-01"), y=20, label = paste("t0max=",round(thourlyregion[[i]]$t0maxyears[1],digits=2)), fontface="italic")
#scale_x_datetime(labels=date_format("%Y-%m"), breaks=("1 year"), minor_breaks=("1 month"))
)
title1=textGrob(names(tempsumlist)[x], gp=gpar(fontface="bold",cex=1.5)) #adjust main title
marrangeGrob(pl,ncol=2,nrow=3,top=title1) #used to multiplot (like par(mfrow))

```



\pagebreak

# References
Tedeschi R., Baldessari M., Mazzoni V., Trona F. and Angeli G. (2012). Population dynamics of *Cacopsylla melanoneura* (Hemiptera: Psyllidae) in northeast Italy and its role in the apple proliferation epidemiology in apple orchards. - Journal of Economic Entomology 105, 322–328.

\pagebreak

## Session info
```{r, echo=FALSE, eval=FALSE}
devtools::session_info()
```


setting  | value  
-------- | -----------------------------  
version  | R version 3.3.1 (2016-06-21)   
system   | x86_64, darwin13.4.0            
ui       | RStudio (1.0.143)                       
language | (EN)                         
collate  | en_US.UTF-8                  
tz       | Europe/Rome   



package     | * | version | date       | source  
----------- | - | --------|------------|---------------  
assertthat  |   | 0.1     | 2013-12-06 | CRAN (R 3.3.0)
 backports  |   | 1.0.5   | 2017-01-18 | CRAN (R 3.3.2)
 colorspace |   | 1.3-2   | 2016-12-14 | CRAN (R 3.3.2)
 DBI        |   | 0.5-1   | 2016-09-10 | CRAN (R 3.3.0)
 devtools   |   | 1.12.0  | 2016-06-24 | CRAN (R 3.3.0)
 digest     |   | 0.6.12  | 2017-01-27 | CRAN (R 3.3.1)
 dplyr      | * | 0.5.0   | 2016-06-24 | CRAN (R 3.3.0)
 evaluate   |   | 0.10    | 2016-10-11 | cran (@0.10)  
 ggplot2    | * | 2.2.1   | 2016-12-30 | CRAN (R 3.3.2)
 gridExtra  | * | 2.2.1   | 2016-02-29 | CRAN (R 3.3.0)
 gtable     |   | 0.2.0   | 2016-02-26 | CRAN (R 3.3.0)
 htmltools  |   | 0.3.5   | 2016-03-21 | CRAN (R 3.3.0)
 knitr      |   | 1.15.1  | 2016-11-22 | CRAN (R 3.3.2)
 labeling   |   | 0.3     | 2014-08-23 | CRAN (R 3.3.0)
 lattice    |   | 0.20-34 | 2016-09-06 | CRAN (R 3.3.0)
 lazyeval   |   | 0.2.0   | 2016-06-12 | CRAN (R 3.3.0)
 magrittr   |   | 1.5     | 2014-11-22 | CRAN (R 3.3.0)
 MASS       | * | 7.3-45  | 2016-04-21 | CRAN (R 3.3.1)
 memoise    |   | 1.0.0   | 2016-01-29 | CRAN (R 3.3.0)
 munsell    |   | 0.4.3   | 2016-02-13 | CRAN (R 3.3.0)
 plyr       | * | 1.8.4   | 2016-06-08 | CRAN (R 3.3.0)
 qtlcharts  | * | 0.7-8   | 2016-06-28 | CRAN (R 3.3.0)
 R6         |   | 2.2.0   | 2016-10-05 | cran (@2.2.0) 
 Rcpp       |   | 0.12.9  | 2017-01-14 | CRAN (R 3.3.2)
 rmarkdown  |   | 1.3     | 2016-12-21 | CRAN (R 3.3.2)
 rprojroot  |   | 1.2     | 2017-01-16 | CRAN (R 3.3.2)
 rsconnect  |   | 0.7     | 2016-12-21 | CRAN (R 3.3.2)
 scales     | * | 0.4.1   | 2016-11-09 | CRAN (R 3.3.2)
 sp         |   | 1.2-4   | 2016-12-22 | CRAN (R 3.3.2)
 stringi    | * | 1.1.2   | 2016-10-01 | cran (@1.1.2) 
 stringr    |   | 1.1.0   | 2016-08-19 | cran (@1.1.0) 
 tibble     |   | 1.2     | 2016-08-26 | CRAN (R 3.3.0)
 withr      |   | 1.0.2   | 2016-06-20 | CRAN (R 3.3.0)
 yaml       |   | 2.1.14  | 2016-11-12 | cran (@2.1.14)


